EDA of DSRCT Samples

### 2024-10-22 analyses/cell-type-dsrct/exploratory_analysis/01-marker-gene-tumor-classification.Rmd
updated the notebooks
updated the notebook to include plot outputs
updated the reference to include papers and internal reference
added plot outputs

```{r}
# source in helper functions: make_jaccard_matrix() and jaccard() 
jaccard_functions <- file.path(module_base, "scripts", "utils", "jaccard-functions.R")
source(jaccard_functions)
```

@@

```{r, fig.width=8}
for(i in 1:length(umap_df_list)){
  # faceted umap showing a umap panel for each marker gene
  p <- ggplot(umap_df_list[[i]], aes(x = UMAP1, y = UMAP2, color = gene_expression)) +
    geom_point(alpha = 0.8, size = 0.1) +
    facet_wrap(vars(gene_symbol)) +
    scale_color_viridis_c() +
    labs(color = "Log-normalized gene expression") +
    # remove axis numbers and background grid
    scale_x_continuous(labels = NULL, breaks = NULL) +
    scale_y_continuous(labels = NULL, breaks = NULL) +
    theme(
      aspect.ratio = 1,
      legend.position = "bottom",
      axis.title = element_text(size = 9, color = "black"),
      strip.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 8)
    ) +
    guides(colour = guide_colorbar(title.position = "bottom", title.hjust = 0.5)) +
    ggtitle(names(umap_df_list)[i])
  print(p)
}

@@

```{r}
for(i in 1:length(umap_df_list_excluded)) {
  p <- ggplot(umap_df_list_excluded[[i]],
              aes(x = gene_expression, fill = gene_symbol)) +
    geom_density() +
    facet_wrap(vars(gene_symbol), scales = 'free_y') +
    theme(legend.position = "none") + 
    ggtitle(names(umap_df_list_excluded)[i])
  print(p)
}
```

@@

for(i in 1:length(umap_df_list_excluded_scaled)) {
  p <- ggplot(umap_df_list_excluded_scaled[[i]],
              aes(x = transformed_gene_expression, fill = gene_symbol)) +
    geom_density() +
    facet_wrap(vars(gene_symbol), scales = 'free_y') +
    theme(legend.position = "none") + 
    ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)
}
```

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_exp)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_viridis_c() +
    ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)
}

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_classification)) +
  geom_point(size = 0.5, alpha = 1) +
  ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)
}

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_exp)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_viridis_c() +
    ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)
}

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_classification)) +
  geom_point(size = 0.5, alpha = 1) +
  ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)
}

@@

## Save outputs

```{r}
# get an RDS of the processed data
saveRDS(sce_file_list, '../results/SCPCP000013_sce_file_list.rds')
```

>>>>

@@

```{r, fig.width=8}
for(i in 1:length(umap_df_list)){
  # faceted umap showing a umap panel for each marker gene
  p <- ggplot(umap_df_list[[i]], aes(x = UMAP1, y = UMAP2, color = gene_expression)) +
    geom_point(alpha = 0.8, size = 0.1) +
    facet_wrap(vars(gene_symbol)) +
    scale_color_viridis_c() +
    labs(color = "Log-normalized gene expression") +
    # remove axis numbers and background grid
    scale_x_continuous(labels = NULL, breaks = NULL) +
    scale_y_continuous(labels = NULL, breaks = NULL) +
    theme(
      aspect.ratio = 1,
      legend.position = "bottom",
      axis.title = element_text(size = 9, color = "black"),
      strip.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 8)
    ) +
    guides(colour = guide_colorbar(title.position = "bottom", title.hjust = 0.5)) +
    ggtitle(names(umap_df_list)[i])
  print(p)

  #save plots
  ggsave(
    filename = paste0(
      '../plots/marker_expression_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

```{r}
for(i in 1:length(umap_df_list_excluded)) {
  p <- ggplot(umap_df_list_excluded[[i]],
              aes(x = gene_expression, fill = gene_symbol)) +
    geom_density() +
    facet_wrap(vars(gene_symbol), scales = 'free_y') +
    theme(legend.position = "none") + 
    ggtitle(names(umap_df_list_excluded)[i])
  print(p)

  #save plots
  ggsave(
    filename = paste0(
      '../plots/marker_distribution_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}
```

@@

for(i in 1:length(umap_df_list_excluded_scaled)) {
  p <- ggplot(umap_df_list_excluded_scaled[[i]],
              aes(x = transformed_gene_expression, fill = gene_symbol)) +
    geom_density() +
    facet_wrap(vars(gene_symbol), scales = 'free_y') +
    theme(legend.position = "none") + 
    ggtitle(names(umap_df_list_excluded_scaled)[i])

  print(p)

  #save plots
  ggsave(
    filename = paste0(
      '../plots/marker_distribution_zscaled_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )

}
```

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_exp)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_viridis_c() +
    ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)

  #save plots
  ggsave(
    filename = paste0(
      '../plots/UMAP_marker_expression_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_classification)) +
  geom_point(size = 0.5, alpha = 1) +
  ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)

  #save plots
  ggsave(
    filename = paste0(
      '../plots/UMAP_classification_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_exp)) +
  geom_point(size = 0.5, alpha = 0.5) +
  scale_color_viridis_c() +
    ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)

  #save plots
  ggsave(
    filename = paste0('../plots/UMAP_marker_expression_cluster_', names(umap_df_list)[i], '.png'),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

for(i in 1:length(marker_sum_exp)) {
  p <- ggplot(marker_sum_exp[[i]], aes(x = UMAP1, y = UMAP2, color = sum_classification)) +
  geom_point(size = 0.5, alpha = 1) +
  ggtitle(names(umap_df_list_excluded_scaled)[i])
  print(p)

  #save plots
  ggsave(
    filename = paste0(
      '../plots/UMAP_classification_cluster_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )

}

@@

## Save outputs 
```{r}
# get an RDS of the processed data
saveRDS(sce_file_list, '../results/SCPCP000013_sce_file_list.rds')
```
